Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-03-28 PandaOverview.tex
Iterate and move related work

>>>>

The Production and Distributed Analysis (PanDA) software is a Workload
Management System (WMS) developed to support the operations of the LHC ATLAS
experiment. PanDA is specifically design to manage the execution of distributed
workloads and workflows via pilots.

WMS for distributed executions are a type of middleware, designed to coordinate
specific activities about resources and workloads: discovering and selecting
resources, submitting tasks of workloads and workflows, and monitoring the
execution of those tasks~\cite{marco2009glite}. Pilot is an abstraction that
enables multi-level scheduling by decoupling resource acquisition from workload
scheduling~\cite{turilli2015comprehensive}. Pilots are implemented by resource
placeholders: a request for a certain amount of resources is scheduled on a site
(e.g., by submitting a job or requesting a virtual machine) and once the
resources are acquired, tasks are scheduled directly to the pilot, not to site's
scheduler.

WMS like PanDA implement the pilot abstraction: WMS control the acquisition of
resources by means of pilots and then manage the execution of workloads or
workflows' tasks on those pilots. WMS centralize the management of the execution
process, offering a unified interface to the application layer. Pilots expose a
uniform scheduling interface and execution environment, abstracting the
differences among the heterogeneous resources of multiple sites. Pilot-enabled
WMS are particularly relevant for LHC experiments, where millions of tasks are
executed across multiple sites every month, analyzing and producing petabytes of
data.

### 2017-03-28 bckground.tex
Iterate and move related work

\jhanote{In this section we should talk less about PanDA and more about other
WMS -- highlighting the fact that (i) there isn't a comprehensive systems
(design, architecture and execution properties) discussion of other WMS so
difficult to compare}\mtnote{What dimension(s) should we use to argue that WMS are difficult to compare?}
\jhanote{(ii) real comparison should be to the Gliden-in WMS ecosystem and use
this as opportunity to narrate the reasons for the divergence.}\mtnote{From our
meetings with Sergey: Before Run 1, ATLAS management become uncomfortable with
the pace at which other WMS were being developed, loosing trust in their ability
to deliver production-grade capabilities. ATLAS USA decided to go solo and
developed their own solution. PanDA proven to be able to deliver the
usual pilot-based WMS capabilities at the right scale. For that, it was chosen
over other solutions.}
\jhanote{Also might this section be moved towards the rear, as this is  a
practice paper and not a traditional research paper?}\mtnote{Fine with me.}

The Production and Distributed Analysis (PanDA) software is a Workload
Management System (WMS) developed to support the operations of the LHC ATLAS
experiment. PanDA is specifically design to manage the execution of distributed
workloads and workflows via pilots.

WMS for distributed executions are a type of middleware, designed to coordinate
specific activities about resources and workloads: discovering and selecting
resources, submitting tasks of workloads and workflows, and monitoring the
execution of those tasks~\cite{marco2009glite}. Pilot is an abstraction that
enables multi-level scheduling by decoupling resource acquisition from workload
scheduling~\cite{turilli2015comprehensive}. Pilots are implemented by resource
placeholders: a request for a certain amount of resources is scheduled on a site
(e.g., by submitting a job or requesting a virtual machine) and once the
resources are acquired, tasks are scheduled directly to the pilot, not to site's
scheduler.

WMS like PanDA implement the pilot abstraction: WMS control the acquisition of
resources by means of pilots and then manage the execution of workloads or
workflows' tasks on those pilots. WMS centralize the management of the execution
process, offering a unified interface to the application layer. Pilots expose a
uniform scheduling interface and execution environment, abstracting the
differences among the heterogeneous resources of multiple sites. Pilot-enabled
WMS are particularly relevant for LHC experiments, where millions of tasks are
executed across multiple sites every month, analyzing and producing petabytes of
data.

@@

Currently, the implementation of PanDA WMS has at least two distinguishing
features: the scale at which it operates distributed computing, and
network-aware scheduling of tasks to sites\mtnote{PanDA team: please let me know
whether more unique (i.e., not present in other LHC WMS) features of PanDA
implementation should be indicated}. PanDA concurrently supports distributed
computations on up to 240,000 cores, submitting more than 365 million jobs a
year across 100 sites, and processing up to XXXXPB of data distributed worldwide
on XX Tier 1-3 facilities. As for 2017, PanDA serves several thousand users,
including XX production groups. The scale of this operation is unprecedented,
making PanDA the WMS that manages the largest computational campaign in the
world.

PanDA was initially implemented to support computations across an unreliable and
relatively expensive networking infrastructure. Accordingly, PanDA assumed a
rigid model of data placement based on replication and long-term storage of data
on the Grid sites. This model proven too wasteful, requiring storage space for
unused and obsolete data. PanDA implementation was evolved to benefit from the
development of the internet and of the networking infrastructure for scientific
research. Today, PanDA constantly monitors network throughput and latency among
its management components, Grid sites and data facilities. PanDA uses this
information to schedule tasks on sites, depending on the estimated time required
to download and/or replicate input data and to stage out output data.

The dynamic management of data replication and download is an example of a
generalized paradigm shift that distinguishes the implementation of PanDA from
other WMS. New components are being implemented to support: (i) dynamic sizing
of input dataset based on resource capabilities; (ii) dynamic sizing of the
resources held by pilots both in terms of number of cores and duration of the
pilot; and (iii) further abstraction of different types of resources to
reconcile high-throughput and high-performance computing paradigms.

% \subsection{Terminology}
%
% \mtnote{To be moved to the end of the introduction.}

% Each task may have an arbitrary number of properties like number of cores,
% executables, or input/output files. Tasks may have precedence interrelations,
% depending on their data dependences or any other type of dependency mandated
% by the application algorithms. Tasks with precedence relations have to be
% executed serially; otherwise, tasks can be executed concurrently. A workload
% is defined as a set of tasks that can be executed concurrently. As such, a
% workflow can be composed of a set of workloads.

@@

% Notes for refactoring:
% \begin{itemize}
%     \item Naming and nomenclature is a contribution
%     \item 1, 1.5 pages.
% \end{itemize}

>>>>

\jhanote{In this section we should talk less about PanDA and more about other
WMS -- highlighting the fact that (i) there isn't a comprehensive systems
(design, architecture and execution properties) discussion of other WMS so
difficult to compare}\mtnote{Better?}
\jhanote{(ii) real comparison should be to the Gliden-in WMS ecosystem and use
this as opportunity to narrate the reasons for the divergence.}\mtnote{From our
meetings with Sergey: Before Run 1, ATLAS management become uncomfortable with
the pace at which other WMS were being developed, loosing trust in their ability
to deliver production-grade capabilities. ATLAS USA decided to go solo and
developed their own solution. PanDA proven to be able to deliver the
usual pilot-based WMS capabilities at the right scale. For that, it was chosen
over other solutions.}
\jhanote{Also might this section be moved towards the rear, as this is  a
practice paper and not a traditional research paper?}\mtnote{Done.}

@@

% -----------------------------------------------------------------------------
% NOTES ON TASK, JOBS, AND WORKFLOW TERMINOLOGY
% -----------------------------------------------------------------------------
% Currently, the implementation of PanDA WMS has at least two distinguishing
% features: the scale at which it operates distributed computing, and
% network-aware scheduling of tasks to sites\mtnote{PanDA team: please let me
% know whether more unique (i.e., not present in other LHC WMS) features of
% PanDA implementation should be indicated}. PanDA concurrently supports
% distributed computations on up to 240,000 cores, submitting more than 365
% million jobs a year across 100 sites, and processing up to XXXXPB of data
% distributed worldwide on XX Tier 1-3 facilities. As for 2017, PanDA serves
% several thousand users, including XX production groups. The scale of this
% operation is unprecedented, making PanDA the WMS that manages the largest
% computational campaign in the world.

% PanDA was initially implemented to support computations across an unreliable
% and relatively expensive networking infrastructure. Accordingly, PanDA
% assumed a rigid model of data placement based on replication and long-term
% storage of data on the Grid sites. This model proven too wasteful, requiring
% storage space for unused and obsolete data. PanDA implementation was evolved
% to benefit from the development of the internet and of the networking
% infrastructure for scientific research. Today, PanDA constantly monitors
% network throughput and latency among its management components, Grid sites and
% data facilities. PanDA uses this information to schedule tasks on sites,
% depending on the estimated time required to download and/or replicate input
% data and to stage out output data.

% The dynamic management of data replication and download is an example of a
% generalized paradigm shift that distinguishes the implementation of PanDA from
% other WMS. New components are being implemented to support: (i) dynamic sizing
% of input dataset based on resource capabilities; (ii) dynamic sizing of the
% resources held by pilots both in terms of number of cores and duration of the
% pilot; and (iii) further abstraction of different types of resources to
% reconcile high-throughput and high-performance computing paradigms.
% -----------------------------------------------------------------------------

% -----------------------------------------------------------------------------
% NOTES ON TASK, JOBS, AND WORKFLOW TERMINOLOGY
% -----------------------------------------------------------------------------
% Each task may have an arbitrary number of properties like number of cores,
% executables, or input/output files. Tasks may have precedence interrelations,
% depending on their data dependences or any other type of dependency mandated
% by the application algorithms. Tasks with precedence relations have to be
% executed serially; otherwise, tasks can be executed concurrently. A workload
% is defined as a set of tasks that can be executed concurrently. As such, a
% workflow can be composed of a set of workloads.

@@

% Notes for refactoring:
% \begin{itemize}
%     \item Naming and nomenclature is a contribution
%     \item 1, 1.5 pages.
% \end{itemize}
% -----------------------------------------------------------------------------
